Parsing mPFC: A Joint Meta-Analytic and Graph Theoretic Approach

### 2018-09-06 Manuscript.Rmd
Working on results and fixed some references

---
title: 'Parsing mPFC: A Joint Meta-Analytic and Graph Theoretic Approach'
author: "Claudio Toro-Serey & Joseph T. McGuire"
output:
  pdf_document: default
  html_document: default
  word_document: 
    reference_docx: Style_reference.docx
bibliography: mPFC_paper.bib
header-includes:
    - \usepackage{setspace}\doublespacing
---

@@

``` {r For HCP data, fig.align="center", echo=FALSE, message = FALSE}
setwd('./For_HCP')
temp <- list.files(pattern = '*Binary*')
FB <- lapply(temp, read.table)
Summed <- do.call(cbind, lapply(FB, "[[", 'V1')) - 1
FB_Means <- rowMeans(Summed)

@@

In order to quantify the intrinsic connectivity of these ROIs in the brain, we acquired resting-state fMRI data from the Human Connectome Project (HCP) [@VanEssen2012] Q6 release (n = 80, randomly sampled from the total pool of 469 available subjects). Each subject's data was acquired over two days at Washington University in St. Louis on a Siemens CONNECTOM Skyra MRI scanner (Siemens, Erlangen, Germany). Four resting state sessions (TR = 0.720 s, TE = 33.1 ms, FA = 52°, multiband factor = 8, 72 slices, 2 mm isotropic voxels) were each comprised of 1200 TRs for a total scan time of 14 min 33 s, with left-right and left-right phase encoding implemented on each day. We used this scanning structure to estimate each subject's topology based on session, day, and overall data (4800 TRs).

@@

To establish each subject's network, we selected all the vertices contained within the ROIs (n = 5,319 per subject) and computed the Pearson correlation of the time series for every pair of vertices. This produced a weighted network for each subject, where surface vertices were the nodes, and edges the correlations among all of them. Next, we applied Fisher's r to z transformation, and took the exponential of the correlations so that all weights were positive while maintaining the shape of the original correlation distribution, and retained the edges whose adjusted p-values remained significant after correction for multiple tests (FDR < 0.05). As part of the evaluation step, this procedure (and the community detection outlined below) was applied at four levels: 1) each session separately (1200 TRs each); 2) the concatenated timeseries from each pair of daily sessions (2400 TRs total); 3) each subject's full dataset (4800 TRs total); and 4) on each step of a sliding window analysis (see Evaluation for more details).

@@

```{r Meta-analysis ROIs, out.width="60%", fig.align = "center", fig.cap= "Figure 1. Template atlas from Glasser et al. [-@Glasser2016] (top), as well as the common and unique brain regions associated with DMN and subjective value (bottom). Overlapping ROIs (in blue) were defined as the common brain regions that were significantly present in their respective literatures, while a single DMN-specific region (7m, in red) was identified from a permutation analysis comparing these literatures."}
include_graphics("./Images/ROIS2.png")
```

@@

``` {r Load Summary Data}
# Load data
setwd('./Summary/')
temp <- list.files()
subjList <- sapply(temp, substring, first=1, last=6)
Summaries <- lapply(temp, read.csv)
nSubj <- length(Summaries)
```

@@

With these 16 ROIs (15 overlapping, 1 DMN-specific), we estimated each individual's DMN coverage using the whole time series (4800 TRs). Each subject's network (n = `r nSubj`) was constructed by performing pairwise correlations among the time series of all the surface vertices contained within the ROIs (`r nrow(Summaries[[1]])` total vertices). After correcting for multiple tests, we performed spectral partitioning and modularity on these networks to identify cohesive functional communities denoting DMN and non-DMN areas. On average, networks had high densities (i.e. highly connected) regardless of the amount of data to produce them (mean = `r round(mdAll, digits = 2)`, SD = `r sddAll`). This degree of density is important, as these methods are known to face challenges when computed on sparse networks [@Fortunato2016]. Figure 2 shows a stereotypical subdivision of the ROIs for a single subject (100307) using spectral partitioning.

@@

This first step just gives an idea of the level of agreement within and between subjects. As a first measure, we use the adjusted rand index (ARI) to estimate the level of agreement between spectral partitioning and modularity. RI measures the proportion of pairs of nodes in a network that were either clustered together or separately in both partitions, over all possible affiliation changes (e.g. clustered together in partition 1, but separately in 2). It is worth noting that the adjusted rand index has well known limitations (Fortunato & Hric, 2016). To compensate for this, we also performed all the upcoming analyses with the variation of information index (VI). This distance measure denotes how much information is gained by adding the second partition (thus, lower values are better), and is known for its robustness. Both RI and VI analyses agreed in the results they yielded, but we only report RI due to its easier interpretability. This first plot shows the distribution of RI values between modularity and spectral partitioning for each subject.

@@

``` {r Partition comparison: all 2, fig.align="center", fig.width = 5, fig.height = 5, fig.cap = "Figure 4. Similarity matrix showing ARI values among all subjects for PCC (lower triangle) and mPFC (upper triangle) separately. PCC is well aligned across individuals, while mPFC shows clear topographical heterogeneity.", echo=FALSE, message = FALSE, cache = TRUE}
# Across subjects
# Note: the upper and lower triangles can totally be combined before plotting, instead of adding one corrplot over another. Might work on that eventually.
RI_all_between <- comparePartitions(Data = Summaries, Index = "RI", nSubjects = length(Summaries), subjNames = subjList)
posteriorIndx <- Summaries[[1]]$y < 0
tempComm_PCC <- lapply(Summaries, "[", i = posteriorIndx, j =)
tempComm_PFC <- lapply(Summaries, "[", i = !posteriorIndx, j =)
RI_PFC <- comparePartitions(Data = tempComm_PFC, Index = "RI", nSubjects = length(Summaries), subjNames = subjList)
RI_PCC <- comparePartitions(Data = tempComm_PCC, Index = "RI", nSubjects = length(Summaries), subjNames = subjList)
# par(mfrow = c(1,2))
# corrplot(RI_all_between,
#          is.corr = F, 
#          method = "color", 
#          tl.pos = "n", 
#          tl.col = "black", 
#          cl.lim = c(0,1),
#          cl.pos = "b",
#          cl.length = 3)
# mtext("Subjects", side = 2, line = -4.5, cex = 1.5)
# mtext("Subjects", side = 3, line = 4.5, cex = 1.5)
corrplot(RI_PCC,
         is.corr = F, 
         method = "color",
         type = "lower", 
         tl.pos = "n", 
         tl.col = "black", 
         cl.lim = c(0,1),
         cl.length = 3,
         cl.pos = "b")
corrplot(RI_PFC, 
         is.corr = F, 
         method = "color", 
         type = "upper", 
         tl.pos = "n", 
         tl.col = "black",
         cl.pos = "n", 
         add = T)

@@

>>>>

---
title: 'Parsing mPFC: A Joint Meta-Analytic and Graph Theoretic Approach'
author: "Claudio Toro-Serey & Joseph T. McGuire"
output:
  word_document: 
    reference_docx: Style_reference.docx
  pdf_document: default
  html_document: default
bibliography: mPFC_paper.bib
header-includes:
    - \usepackage{setspace}\doublespacing
---

@@

``` {r For HCP data, fig.align="center", echo=FALSE, message = FALSE}
setwd('./For_HCP')
temp <- list.files(pattern = '*Binary*')
FB <- lapply(temp, read.table)
Summed <- do.call(cbind, lapply(FB, "[[", 'V1')) - 1
FB_Means <- rowMeans(Summed)
nSubj <- length(temp)
rm(temp)
```

@@

In order to quantify the intrinsic connectivity of these ROIs in the brain, we acquired resting-state fMRI data from the Human Connectome Project (HCP) [@VanEssen2012] Q6 release (n = `r nSubj`, randomly sampled from the total pool of 469 available subjects). Each subject's data was acquired over two days at Washington University in St. Louis on a Siemens CONNECTOM Skyra MRI scanner (Siemens, Erlangen, Germany). Four resting state sessions (TR = 0.720 s, TE = 33.1 ms, FA = 52°, multiband factor = 8, 72 slices, 2 mm isotropic voxels) were each comprised of 1200 TRs for a total scan time of 14 min 33 s, with left-right and left-right phase encoding implemented on each day. We used this scanning structure to estimate each subject's topology based on session, day, and overall data (4800 TRs).

@@

To establish each subject's network, we selected all the vertices contained within the ROIs (n = 5,081 per subject) and computed the Pearson correlation of the time series for every pair of vertices. This produced a weighted network for each subject, where surface vertices were the nodes, and edges the correlations among all of them. Next, we applied Fisher's r to z transformation, and took the exponential of these values so that all weights were positive while maintaining the shape of the original correlation distribution, and retained the edges whose adjusted p-values remained significant after correction for multiple tests (FDR < 0.05). As part of the evaluation step, this procedure (and the community detection outlined below) was applied at four levels: 1) each session separately (1200 TRs each); 2) the concatenated timeseries from each pair of daily sessions (2400 TRs total); 3) each subject's full dataset (4800 TRs total); and 4) on each step of a sliding window analysis (see Evaluation for more details).

@@

```{r Meta-analysis ROIs, out.width="60%", fig.align = "center", fig.cap= "Figure 1. Template atlas from Glasser et al. (2016) (top), as well as the common and unique brain regions associated with DMN and subjective value (bottom). Overlapping ROIs (in blue) were defined as the common brain regions that were significantly present in their respective literatures, while a single DMN-specific region (7m, in red) was identified from a permutation analysis comparing these literatures."}
include_graphics("./Images/ROIS2.png")
```

@@

``` {r Load Summary Data}
# Load data
setwd('./Summary/')
temp <- list.files()
subjList <- sapply(temp, substring, first=1, last=6)
Summaries <- lapply(temp, read.csv)
```

@@

With these 16 ROIs (15 overlapping, 1 DMN-specific), we estimated each individual's DMN coverage using the whole time series (4800 total TRs from four scanning sessions over two days). Each subject's network (n = `r nSubj`) was constructed by performing pairwise correlations among the time series of all the surface vertices contained within the ROIs (`r nrow(Summaries[[1]])` total vertices). After correcting for multiple testing, we performed spectral partitioning and modularity on these networks to identify cohesive functional communities denoting DMN and non-DMN areas. On average, networks had high densities (i.e. highly connected) regardless of the amount of data to produce them (mean = `r round(mdAll, digits = 2)`, SD = `r sddAll`). This degree of density is important, as these methods are known to face challenges when computed on sparse networks [@Fortunato2016].

Figure 2 shows a stereotypical subdivision of the ROIs for a single subject (100307) using spectral partitioning. This method dissects a network by the positive and negative values derived from the Fiedler vector (the eigenvector related to the second-to-lowest eigenvalue of the network's Laplacian matrix, see Methods), which we used to binarize the partition. Since this is a data-driven approach, there is no a priori labeling for these communities. We thus defined the DMN community as the one containing the one DMN-specific area from the meta-analysis (7m), and ensured that the binary labeling of the vertices was consistent across subjects. In general, the resulting pattern shows large DMN coverage in PCC, with non-DMN vertices populating its anterior segment. mPFC, on the other hand, is mostly populated by DMN vertices in its ventral-anterior and dorsal-anterior areas, with a persistent non-DMN pocket between them. This non-DMN section extends posteriorly onto pregenual cinglate cortex. Note that vmPFC does not encompass area a25, which is often related to negative emotional valuation [@Joyce2018]. Interstingly, this topographical combination of dorsal and ventral mPFC, with extensive PCC coverage, matches previous meta-analytic patterns more closely attributed to social cognition than SV [@Hiser2018].

@@

Before determining the degree of generazibility of this topological pattern across individuals, we examined the validity of these partitions by comparing them to patterns derived with a modularity algorithm [@Clauset2004]. Modularity seeks to find the pattern of communities that maximizes a quality function that compares the connectivity strength of any two nodes with a null probability model. Since modularity is agnostic to the desired number of clusters, it was capable of finding more than two communities in our datset. Thus, if both methods agreed in the number and topological organization of the current communities, it would add confidence to our results. We quantified the cross-method agreement with the adjusted rand index (ARI). ARI measures the proportion of pairs of nodes in a network that were either clustered together or separately in both partitions, over all possible affiliation changes (e.g. clustered together in partition 1, but separately in 2), and has the benefit of being agnostic to labeling schemes [@Hubert1985]. Figure 3 shows the distribution of ARI values across subjects. **It is worth noting that the adjusted rand index has well known limitations (Fortunato & Hric, 2016). To compensate for this, we also performed all the upcoming analyses with the variation of information index (VI). This distance measure denotes how much information is gained by adding the second partition (thus, lower values are better), and is known for its robustness. Both RI and VI analyses agreed in the results they yielded, but we only report RI due to its easier interpretability. This first plot shows the distribution of RI values between modularity and spectral partitioning for each subject.**

@@

``` {r Partition comparison: all 2, fig.align="center", fig.width = 4, fig.height = 4, fig.cap = "Figure 4. Similarity matrix showing ARI values among all subjects for PCC (lower triangle) and mPFC (upper triangle) separately. PCC is well aligned across individuals, while mPFC shows clear topographical heterogeneity.", echo=FALSE, message = FALSE, cache = TRUE}
# Across subjects
# Note: the upper and lower triangles can totally be combined before plotting, instead of adding one corrplot over another. Might work on that eventually.
RI_all_between <- comparePartitions(Data = Summaries, Index = "RI", nSubjects = length(Summaries), subjNames = subjList)
posteriorIndx <- Summaries[[1]]$y < 0
tempComm_PCC <- lapply(Summaries, "[", i = posteriorIndx, j =)
tempComm_PFC <- lapply(Summaries, "[", i = !posteriorIndx, j =)
RI_PFC <- comparePartitions(Data = tempComm_PFC, Index = "RI", nSubjects = length(Summaries), subjNames = subjList)
RI_PCC <- comparePartitions(Data = tempComm_PCC, Index = "RI", nSubjects = length(Summaries), subjNames = subjList)
# par(mfrow = c(1,2))
# corrplot(RI_all_between,
#          is.corr = F, 
#          method = "color", 
#          tl.pos = "n", 
#          tl.col = "black", 
#          cl.lim = c(0,1),
#          cl.pos = "b",
#          cl.length = 3)
# mtext("Subjects", side = 2, line = -4.5, cex = 1.5)
# mtext("Subjects", side = 3, line = 4.5, cex = 1.5)
corrplot(RI_PCC,
         is.corr = F, 
         method = "color",
         type = "lower", 
         tl.pos = "n", 
         tl.col = "black", 
         cl.lim = c(0,1),
         cl.length = 3,
         cl.pos = "b")
corrplot(RI_PFC, 
         is.corr = F, 
         method = "color", 
         type = "upper", 
         tl.pos = "n", 
         tl.col = "black",
         cl.pos = "n", 
         add = T)

@@

It's also worth mentioning that PCC is largely covered by DMN. Then why does this area overlap with SV? I think the Mueller paper mentions that PCC is an important hub of the brain, and perhaps subserves many cognitive functions.

And write a portion on the lack of subcortical reward-related regions (i.e. striatum). Say something like 'incorporating the ventral striatum did notrefine these partitions through either method. This could be due to the lack of signal modulation of this area at rest, and should be examined through behavioral manipulations of value in the future. Instead, here we focus on capturing the intrinsic functional heterogeneity of mPFC as a way to pave the path to refined analyses aimed at disentangling DMN and SV'.

Talk about how there are other community detection algorithms, and that modularity and SP over and underfit, respectively. However, SP eigenmaps have been found to provide relevant functional information. In addition, both modularity and SP rely on connectivity strength, and thus is amenable to the correlative nature of functional connectivity.

Add a portion showing that summing up the binarized maps produces a map similar to the one produced by Margulies et al. and others.
